---
title: "Project 4"
author: "Matthew Leong, Mervyn Jonathon, Luke Bravo, Joe Niehaus"
date: "4/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
#loading in libraries
library(gurobi)
library(ggplot2)
library(tidyverse)
# import data
data <- read.csv("price_demand_data.csv") 
```


### Fitting Linear Regression to model demand (1+2)

```{r linreg,echo = TRUE}
#do a linear regression using R
lin_model = lm(demand ~ price, data)
summary(lin_model)
#look at summary
#coefficient1 is intercept
#coefficient2 is price coefficient
#residuals - refer to this if you need a refresher: https://www.statisticshowto.com/residual/

# printing cost
c = 0.5
# rush print cost
g = 0.75
# disposal fee
t = 0.15
# price
p = 1

# Get demand residual points of the linear regression and store it
data_resid = data %>% mutate(residuals = residuals(lin_model))

# Add the residuals to the data to get demand at price 1 
#(intuitively residuals are unexplained errors, add them to the model to get 100% explanation from the model.)
#If you want to check this, I suggest changing p to a price in the demand data. You will get the demand at that price.
demand_p1 = data_resid %>% mutate(pred_demandp1 = lin_model$coefficients[1] + lin_model$coefficients[2]*p + data_resid[,"residuals"])
```

### Finding Optimal Quantity to Produce when price is 1 (3)

This problem is similar to the one we did in class but with some key differences. Namely, we have two conditions. I detail the first as when we produce too much and the second as when we produce too little both relative to the optimal demand. A note about this condition is that since rush printing costs are still less than the revenue we get from printing them, we always choose to rush print. As a reminder for the following c is printing cost, g is rushed printing cost, t is disposal fee, and p is price.
$$
pD_i-qc-t(q-D_i) \, if \, q > D_i \\
pD_i-qc-g(D_i-q) \, if \, q \leq D_i
$$
Now reformulate the above into constraints just like the newsvendor problem, A matrix should have anything related to h and q, right hand side should have anything related to demand. Regarding the case where demand exceeds quantity first:
$$
h_i \leq pD_i-qc-t(q-D_i) \\
h_i + qc +t(q-D_i) \leq pD_i \\
h_i + q(c+t) \leq (p+t)D_i \\
$$
If you're confused about why t is positive for demand. Think of it as canceling out the disposal fee in the quantity side. tD cancels part of tq essentially. There is still a disposal fee but now it correctly only takes into account the quantity that is actually disposed (q-D)t.
$$
h_i \leq pD_i - qc - g(D_i-q) \\
h_i +qc-qg \leq pD_i - gD_i \\
h_i +q(c-g) \leq (p-g)D_i
$$
Again for the cost of rushed newspapers. the above seems weird but think of it like this: p-g is the penalty assigned to newspapers. c-g offsets this penalty for all q prepared before hand. The remaining (D-q) then eat up the g penalty. So mathematically, q newspapers cost c and D-q newspapers cost g.
  
Regardless, these two formulations go in the constraint and rhs.

```{r LP,echo = TRUE}
#this is pretty similar to our newsvendor problem. See lecture on April 1st.
#get demand for price 1
demand = demand_p1[,"pred_demandp1"]
# number of days of demand data, if you check data, there are 99 rows which this matches
nd = length(demand) 

# decision variables are (q, h1, h2, ..., h99)
#objective is the same as in class. Take the average of all the days. Letting h be profit for each day.
obj = c(0,rep(1/nd,nd))
lb = c(0,rep(-Inf,nd)) # quantity printed needs to be non-negative, but profit on a given day could possibly be negative

#number of constraints is number of days*2 ()
A = matrix(0,2*nd,nd+1)
rhs = rep(0,2*nd)
dir = rep('<=',2*nd)


for(r in 1:nd){ #for every day create two constraints
  #This is the constraint where q > D
  #h_i + q(c+t) \leq (p+t)D_i 
  A[2*r-1,c(1,r+1)] = c(c+t,1) # q first, 1 for h_i
  rhs[2*r-1] = (p+t)*demand[r] #represents the demand part
  
  #This is the constraint where q < D
  #h_i +q(c-g) \leq (p-g)D_i
  A[2*r,c(1,r+1)] = c(c-g,1) # q first, 1 for h_i
  rhs[2*r] = (p-g)*demand[r]
}

nv.model = list()
nv.model$modelsense = 'max'
nv.model$obj = obj
nv.model$A = A
nv.model$rhs = rhs
nv.model$sense = dir
nv.model$lb = lb

pars = list()
pars$outputflag = 0

nv.sol = gurobi(nv.model,params=pars)
nv.sol$x[1] # how many we should produce
nv.sol$objval # expectation of profits given that many produced - ish...
```

### Quadratic Problem (4)
I'm doing it QP since that means less matrix multiplication overall. The suggestion is having h be the negative cost on each day. The quadratic part would be the average revenue. General form:  
$$
max \, x^{T} Q x + c^{T}x
$$
How does letting price impact demand change the problem? Recall step 1 and 2. 
$$
p*D_i = p(\beta_0+\beta_1*p+\epsilon)
$$
aka
```{r}
lin_model$coefficients[1] + lin_model$coefficients[2]*p + data_resid[,"residuals"]
```

In this problem, we always strive to hit $pD_i$ thus, we can consider average revenue to be then:
$$
\frac{\beta_1p^2+(\beta_0 +\epsilon)p}{numberofdays}*numberofdays
$$
Negative cost on the other hand is the hs. Let us then consider the cases:
$$
-qc-t(q-D_i) \, if \, q > D_i \\
-qc-g(D_i-q) \, if \, q \leq D_i
$$
Need to convert D to the appropriate equation:
$$
-qc-t(q -(\beta_0+\beta_1*p+\epsilon)) = -(qc+t(q-\beta_0-\beta_1*p-\epsilon)) \\
-qc-g((\beta_0+\beta_1*p+\epsilon)-q) = -(qc+g(\beta_0+\beta_1*p+\epsilon-q)) 
$$

So bringing that together we have:
$$
Average Revenue - Average Cost each day \\
\frac{\beta_1p^2+\beta_0 p+\epsilon p}{n}*n -(-h_i)
$$
An interesting property if you recall is that the sum of residuals is actually 0. (if you don't believe me you can check: sum(data_resid[,"residuals"])) Average revenue thus is just the two coefficients of the linear regression model multiplied by price.
How to formulate? Decision variables are:
$$
x = [p,q,h_{11},h_{12},...]
$$

New let h be the cost. 

First the more quantity constraints
$$
h_i \leq -(qc+t(q-\beta_0-\beta_1*p-\epsilon)) \\
h_i \leq -qc - -t(q-\beta_0-\beta_1*p-\epsilon) \\
h_i + qc + tq \leq t\beta_0 + t \beta_1*p+t\epsilon \\
h_i + (t+c)q - t\beta_1*p \leq t(\beta_0+\epsilon)
$$


Less quantity constraints:
$$
h_i \leq -(qc+g(\beta_0+\beta_1*p+\epsilon-q)) \\
h_i +qc \leq -g(\beta_0+\beta_1*p+\epsilon-q) \\
h_i +qc-qg + g\beta_1*P \leq -g(\beta_0+\epsilon) 
$$
Interesting thing to note is that only $p^2$ is quadratic meaning that the hint is very applicable. We only need p in the x vector to be quadratic. The rest is linear. 
That means the gurobi objective function should look something like this:
$$
Q = 
\begin{pmatrix}
\beta_1  & 0 & 0 & ..... & 0\\
0 .................... \\
0
\end{pmatrix}
$$
Basically first term is just the coefficient of linear regression divided by number of days. This captures the quadratic part of the model.   
Meanwhile the linear should be:
$$
\beta_0p + 0*q + hpart
$$
hpart is negative costs. Hence the add for h. 
Ok, let's try that in gurobi.


```{r}
#this is pretty similar to our newsvendor problem. See lecture on April 1st.
#get demand for price 1

# number of days of demand data, if you check data, there are 99 rows which this matches
nd = nrow(data) 

# decision variables are (p,q, h1, h2, ..., h99)
# total of 2+nd variables. Using all the data that is 101 variables.
#defining p as the first variable makes things easier.

#defining quadratic term
#in the class notes this is the middle matrix term
#+2 to account for price and quantity.
quad = matrix(0,nd+2,nd+2)
quad[1,1] = lin_model$coefficients[[2]]

#defining linear term
#Since first term is price. Assign the appropriate coefficients to the objective
obj = c(lin_model$coefficients[[1]],0,rep(1/nd,nd))

lb = c(0,0,rep(-Inf,nd)) 
#Maximizing over 

#number of constraints is number of days*2 ()
A = matrix(0,2*nd,nd+2)
rhs = rep(0,2*nd)
dir = rep('<',2*nd)

#friendly reminder that p is first, q is second. 
for(r in 1:nd){ #for every day create two constraints
  #This is the constraint where q > D
  #h_i + (t+c)q - t\beta_1*p \leq t(\beta_0+\epsilon)
  A[2*r-1,c(1,2,r+2)] = c(-t*lin_model$coefficients[2],(c+t),1) 
  rhs[2*r-1] = t*(lin_model$coefficients[1]+data_resid$residuals[[r]])
  
  #This is the constraint where q < D
  #h_i + q(g-c) - gp\beta_1 \leq g(\beta_0 + \epsilon) 
  A[2*r,c(1,2,r+2)] = c(g*(lin_model$coefficients[2]), (c-g) ,1) 
  rhs[2*r] = -g*(lin_model$coefficients[1]+data_resid$residuals[[r]])
}

nvqp.model = list()
nvqp.model$modelsense = 'max'

#this takes in the quadratic part of the objective function
nvqp.model$Q = quad

#this takes the linear part of the objective function
nvqp.model$obj = obj 

#should be fine since linear constraints. 
nvqp.model$A = A
nvqp.model$rhs = rhs
nvqp.model$sense = dir

nvqp.model$lb = lb


params = list()
params$outputflag = 0
qp.nvqp = gurobi(nvqp.model,params=params)

cat('Optimal Price: ', qp.nvqp$x[1],'\n')
cat('Optimal Quantity:  ', qp.nvqp$x[2],'\n')
cat('Optimal Profit: ', qp.nvqp$objval)
```



