---
title: "Project 2"
author: "Matthew Leong, India Lindsay, John Chen, Andrew Han"
date: "3/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gurobi)
library(lpSolve)
library(knitr)
```

## R Markdown


## Read in the data

```{r cars}
# read the data
train_df = read.csv('stocks2019.csv',header = TRUE, row.names = 1)
test_df = read.csv('stocks2020.csv',header = TRUE, row.names = 1)
```

## Calculating returns and correlations of stocks

Returns here are defined as:
$$
\frac{P_1-P_0}{P_0}
$$
where $P_0$ stands for the initial stock price and $P_1$ stands for the stock price at the end of one period which in the provided file is generally a day or 3 days if on a Friday. 
```{r pressure, echo=FALSE}
# let a be the rows of the return matrix. b be the columns.
a = nrow(train_df)
b = ncol(train_df)

#In theory first row should consist of zeros.
#Define as empty 0 matrix at first.
return_matrix = matrix(0,a,b)

#Identify rows as the date
rownames(return_matrix) = rownames(train_df)

#Identify columns as the index fund price and the stocks.
colnames(return_matrix) = colnames(train_df)
```

```{r}
#Iterate columns with i.
#Use special R magic aka operate on all the rows except for the first because that will break it.
#See the LaTex earlier for how the returns are calculated.
for (i in 1:b){
  #Translation from code speak: 
  #returnmatrix for every row except one at column i (starting with 1 which is NDX) =
  #(price at the end of the period - price at the beginning of the period)/(price at the beginning of the period)
  return_matrix[2:a,i] = (train_df[2:a,i] - train_df[1:a-1,i])/train_df[1:a-1,i]
}

#also calculate return matrix for test
c = nrow(test_df)
d = ncol(test_df)
test_return_matrix = matrix(0,c,d)
rownames(test_return_matrix) = rownames(test_df)
colnames(test_return_matrix) = colnames(test_df)
for (i in 1:d){
  test_return_matrix[2:c,i] = (test_df[2:c,i] - test_df[1:c-1,i])/test_df[1:c-1,i]
}
```

```{r}
#getting the correlations is simply just through cor built in function
corr_matrix = cor(return_matrix)

#Remove Index from correlation matrix.
corr_matrix2 = corr_matrix[,-1]
corr_matrix2 = corr_matrix2[-1,]

#Redefine number of columns for stocks.
r = ncol(corr_matrix2)
```


## Stock Selection IP
```{r}
#Choosing the stocks aka stock selection IP
#essentially, get all the correlations for the first part. Next part is to represent the "ys".
#The "ys" should have no bearing on the objective function so we give them values of 0.
obj = c(as.vector(corr_matrix2),rep(0,r))
#How it vectorizes it goes like this: 1,1;1,2;1,3,etc.

#Define constraints. Just how many are there? A lot.
#Need all the correlation entries in the correlation matrix. So r*r or r^2. 
#Have one constraint for the ys aka 1 row
#Have another set of constraints r to signify y_j (a chosen stock)

#Adding all the number of row constraints together: r^2+1+r. Then r^2 for x_ij variables and r for y_j variables.
A = matrix(0,r^2+1+r,r^2+r)

m = 5

#define y variable constraint (no b yet.) 
A[1,(r^2+1):(r^2+r)] = rep(1,r) 

#Row constraint x_ij = 1 sum.r constraints
for (i in 1:r){
  A[i+1,(r*(i-1)+1):(r*i)] = rep(1,r)
}

#r^2 constraints for x_ij \leq y_j. 

#A[102:(102^2+101+1),x_ij variables]
A[(r+2):(r^2+r+1), 1:r^2] = diag(1,r^2)

#Define the y_js
A[(r+2):(r^2+r+1), (r^2+1):(r^2+r)] = matrix(rep(diag(-1,r),r), nrow=r^2, byrow=T)

#In order, m constraint, sum constraints, xij vs yj constraints
b = c(m, rep(1,r), rep(0,r^2))
#it's = for m and sum. < for xij vs yj
sense = c(rep('=',(r+1)),rep('<',r^2))
#They are all binary
vartype = rep('B',r^2+r)

# set up the model
model = list()
model$modelsense = 'max'
model$obj = obj
model$A = A
model$rhs = b
model$sense = sense
model$vtype = vartype

# solve the problem
params = list()
params$outputflag = 0
sol = gurobi(model,params=params)

# optimal allocation:
x = sol$x
```

```{r}
#Get the stocks
stocks = tail(which(x %in% 1),m)
for (i in 1:m) {
  stocks[i] = stocks[i]-r^2
}

#Get the names of the stocks
corr_names = colnames(corr_matrix2)
names = rep(0,5)
for (i in 1:m) {
  names[i] = corr_names[stocks[i]]
}
```



## Weights Problem
$$
min  \sum_{t=1}^T |q_t - \sum_{i=1}^mw_i r_{it} |
$$

#John's explanation
$$
min |q_1-\sum_{i=1}^mw_i*r_{i1}|+|q_2-\sum_{i=1}^mw_i*r_{i2}|...
$$

reformulate to:
$$
min \sum_{i=1}^T y_i
$$
where
$$
y_1 \geq q_1 - \sum_{i=1}^mw_ir_{i1} \\
y_1 \geq \sum_{i=1}^mw_ir_{i1} -q_1
$$

Rows for A:
$$
2*t+1
$$
2t stands for every time.
1 stands for the sum up to 1.

```{r}
t = nrow(return_matrix)
#minimize weights
objective = c(rep(0,m),rep(1,t))


#Rows for A: 2t+1
#2t stands for every time and 1 stands for the sum up to 1 constraint.
A = matrix(0,2*t+1,m+t)
b = rep(0,2*t+1)
sense = rep('',2*t+1)

#Weights sum up to one constraint
A[1,] = c(rep(1,m),rep(0,t))
b[1] = 1
sense[1] = '='

#One half of the absolute value to linear program constraint.
for (i in 1:t){
  A[i+1, m+i] = 1 #set y_i in this constraint to 1
  for (j in 1:m){
    #+1 to compensate for index being in return matrix
    A[i+1,j] = return_matrix[i,stocks[j]+1]
  }
  b[i+1] = return_matrix[i,1]
  sense[i+1] = '>'
}

#Other half of the absolute value to linear program constraint (negative)
for (i in 1:t){
  A[i+t+1, m+i] = 1 #set y_i in this constraint to 1
  for (j in 1:m){
    #+1 to compensate for index being in return matrix
    A[i+t+1,j] = -return_matrix[i,stocks[j]+1]
  }
  b[i+t+1] = -return_matrix[i,1]
  sense[i+t+1] = '>'
}

# set up the model
model = list()
model$modelsense = 'min'
model$obj = objective
model$A = A
model$rhs = b
model$lb = rep(0,m+t)
model$sense = sense

# solve the problem
params = list()
params$outputflag = 0
sol = gurobi(model,params=params)

# optimal allocation:
x = sol$x
```


```{r}
#Store the weights
weights = x[1:5]
#Testing it to 2020
t2 = nrow(test_return_matrix)
compare_sum_2020 = 0
for (t in 1:t2) {
  invest = 0
  for (i in 1:m){
    invest = invest + weights[i]*test_return_matrix[t,stocks[i]+1]
  }
  compare_sum_2020 = compare_sum_2020 + abs(test_return_matrix[t,1]-invest)
}

t = nrow(return_matrix)
compare_sum_2019 = 0
for (j in 1:t) {
  invest = 0
  for (i in 1:m){
    invest = invest + weights[i]*return_matrix[j,stocks[i]+1]
  }
  compare_sum_2019 = compare_sum_2019 + abs(return_matrix[j,1]-invest)
}
```


### Function time
```{r}
#Needs m (number of stocks), correlation matrix
#Returns a list of ids that correspond to the chosen stocks
get_stocks = function(m,corr_mat){
  r = ncol(corr_mat)
  obj = c(as.vector(corr_mat),rep(0,r))
  A = matrix(0,r^2+1+r,r^2+r)
  
  #define y variable constraint (no b yet.) 
  A[1,(r^2+1):(r^2+r)] = rep(1,r) 
  
  #Row constraint x_ij = 1 sum.r constraints
  for (i in 1:r){
    A[i+1,(r*(i-1)+1):(r*i)] = rep(1,r)
  }
  
  #r^2 constraints for x_ij \leq y_j. 

  #A[102:(102^2+101+1),x_ij variables]
  A[(r+2):(r^2+r+1), 1:r^2] = diag(1,r^2)
  
  #Define the y_js
  A[(r+2):(r^2+r+1), (r^2+1):(r^2+r)] = matrix(rep(diag(-1,r),r), nrow=r^2, byrow=T)
  
  #In order, m constraint, sum constraints, xij vs yj constraints
  b = c(m, rep(1,r), rep(0,r^2))
  #it's = for m and sum. < for xij vs yj
  sense = c(rep('=',(r+1)),rep('<',r^2))
  
  #They are all binary
  vartype = rep('B',r^2+r)
  
  # set up the model
  model = list()
  model$modelsense = 'max'
  model$obj = obj
  model$A = A
  model$rhs = b
  model$sense = sense
  model$vtype = vartype
  
  # solve the problem
  params = list()
  params$outputflag = 0
  sol = gurobi(model,params=params)
  
  # optimal allocation:
  x = sol$x
  
  #Get the stocks
  stocks = tail(which(x %in% 1),m)
  for (i in 1:m) {
    stocks[i] = stocks[i]-r^2
  }
  #returns the stocks which has the IDS
  return(stocks)
}
```

```{r}
#Function get_weights accepts a return matrix, number of stocks, and a chosen stock list from get_stocks
#It returns the solution to the integer program.
get_weights_and_returns <- function(return_matrix,m,stocks){
  t = nrow(return_matrix)
  #minimize weights
  objective = c(rep(0,m),rep(1,t))
  
  
  #Rows for A: 2t+1
  #2t stands for every time and 1 stands for the sum up to 1 constraint.
  A = matrix(0,2*t+1,m+t)
  b = rep(0,2*t+1)
  sense = rep('',2*t+1)
  
  #Weights sum up to one constraint
  A[1,] = c(rep(1,m),rep(0,t))
  b[1] = 1
  sense[1] = '='
  
  #One half of the absolute value to linear program constraint.
  for (i in 1:t){
    A[i+1, m+i] = 1 #set y_i in this constraint to 1
    for (j in 1:m){
      #+1 to compensate for index being in return matrix
      A[i+1,j] = return_matrix[i,stocks[j]+1]
    }
    b[i+1] = return_matrix[i,1]
    sense[i+1] = '>'
  }
  
  #Other half of the absolute value to linear program constraint (negative)
  for (i in 1:t){
    A[i+t+1, m+i] = 1 #set y_i in this constraint to 1
    for (j in 1:m){
      #+1 to compensate for index being in return matrix
      A[i+t+1,j] = -return_matrix[i,stocks[j]+1]
    }
    b[i+t+1] = -return_matrix[i,1]
    sense[i+t+1] = '>'
  }
  
  # set up the model
  model = list()
  model$modelsense = 'min'
  model$obj = objective
  model$A = A
  model$rhs = b
  model$lb = rep(0,m+t)
  model$sense = sense
  
  # solve the problem
  params = list()
  params$outputflag = 0
  sol = gurobi(model,params=params)
  
  # optimal allocation:
  x = sol$x
  #Return the solution. x[1:m] are the weights of the stocks. m:(t+m) are the returns for the years. 
  return(x)
}
```

```{r}
#Function get_returns accepts a return matrix, its associated stocks_index_list, and its associated weights.
#It returns the returns compared to the index for the entire year of the return matrix.
get_returns <- function(return_mat,stocks_index_list,weights) {
  t = nrow(return_mat)
  m = length(stocks_index_list)
  return_sum = 0
  for (i in 1:t){
    invest = 0
    for (j in 1:m) {
      invest = invest + weights[j]*return_mat[i,stocks_index_list[j]+1]
    }
  return_sum = return_sum + abs(return_mat[i,1]-invest)
  }
  return(return_sum)
}
```


## 3. Comparing different values of m.

```{r}
compare_sum5_2019 = get_returns(return_matrix,stocks,weights)
compare_sum5_2020 = get_returns(test_return_matrix,stocks,weights)
```

```{r}
ten_stocks = get_stocks(10,corr_matrix2)

ten_stocks_weights_and_returns = get_weights_and_returns(return_matrix,10,ten_stocks)
ten_stocks_weights = ten_stocks_weights_and_returns[1:10]

#This is here to double check. Do sum(ten_Stocks_returns_2019 and compare it to ten_stocks_sum2019)
#The two values should be fairly similar.
ten_stocks_returns_2019 = ten_stocks_weights_and_returns[11:(nrow(return_matrix)+10)]

ten_stocks_sum2019 = get_returns(return_matrix,ten_stocks,ten_stocks_weights)
ten_stocks_sum2020 = get_returns(test_return_matrix,ten_stocks,ten_stocks_weights)
```


```{r}
twen_stocks = get_stocks(20,corr_matrix2)

twen_stocks_weights_and_returns = get_weights_and_returns(return_matrix,20,twen_stocks)
twen_stocks_weights = twen_stocks_weights_and_returns[1:20]
twen_stocks_returns_2019 = twen_stocks_weights_and_returns[21:(nrow(return_matrix)+20)]

twen_stocks_sum2019 = get_returns(return_matrix,twen_stocks,twen_stocks_weights)
twen_stocks_sum2020 = get_returns(test_return_matrix,twen_stocks,twen_stocks_weights)
```

```{r}
thirty_stocks = get_stocks(30,corr_matrix2)

thirty_stocks_weights_and_returns = get_weights_and_returns(return_matrix,30,thirty_stocks)
thirty_stocks_weights = thirty_stocks_weights_and_returns[1:30]
thirty_stocks_returns_3019 = thirty_stocks_weights_and_returns[31:(nrow(return_matrix)+30)]

thirty_stocks_sum2019 = get_returns(return_matrix,thirty_stocks,thirty_stocks_weights)
thirty_stocks_sum2020 = get_returns(test_return_matrix,thirty_stocks,thirty_stocks_weights)
```

```{r}
forty_stocks = get_stocks(40,corr_matrix2)

forty_stocks_weights_and_returns = get_weights_and_returns(return_matrix,40,forty_stocks)
forty_stocks_weights = forty_stocks_weights_and_returns[1:40]
forty_stocks_returns_4019 = forty_stocks_weights_and_returns[41:(nrow(return_matrix)+40)]

forty_stocks_sum2019 = get_returns(return_matrix,forty_stocks,forty_stocks_weights)
forty_stocks_sum2020 = get_returns(test_return_matrix,forty_stocks,forty_stocks_weights)
```

```{r}
fifty_stocks = get_stocks(50,corr_matrix2)

fifty_stocks_weights_and_returns = get_weights_and_returns(return_matrix,50,fifty_stocks)
fifty_stocks_weights = fifty_stocks_weights_and_returns[1:50]
fifty_stocks_returns_5019 = fifty_stocks_weights_and_returns[51:(nrow(return_matrix)+50)]

fifty_stocks_sum2019 = get_returns(return_matrix,fifty_stocks,fifty_stocks_weights)
fifty_stocks_sum2020 = get_returns(test_return_matrix,fifty_stocks,fifty_stocks_weights)
```

```{r}
sixty_stocks = get_stocks(60,corr_matrix2)

sixty_stocks_weights_and_returns = get_weights_and_returns(return_matrix,60,sixty_stocks)
sixty_stocks_weights = sixty_stocks_weights_and_returns[1:60]
sixty_stocks_returns_2019 = sixty_stocks_weights_and_returns[61:(nrow(return_matrix)+60)]

sixty_stocks_sum2019 = get_returns(return_matrix,sixty_stocks,sixty_stocks_weights)
sixty_stocks_sum2020 = get_returns(test_return_matrix,sixty_stocks,sixty_stocks_weights)
```

```{r}
seventy_stocks = get_stocks(70,corr_matrix2)

seventy_stocks_weights_and_returns = get_weights_and_returns(return_matrix,70,seventy_stocks)
seventy_stocks_weights = seventy_stocks_weights_and_returns[1:70]
seventy_stocks_returns_2019 = seventy_stocks_weights_and_returns[71:(nrow(return_matrix)+70)]

seventy_stocks_sum2019 = get_returns(return_matrix,seventy_stocks,seventy_stocks_weights)
seventy_stocks_sum2020 = get_returns(test_return_matrix,seventy_stocks,seventy_stocks_weights)
```

```{r}
eighty_stocks = get_stocks(80,corr_matrix2)

eighty_stocks_weights_and_returns = get_weights_and_returns(return_matrix,80,eighty_stocks)
eighty_stocks_weights = eighty_stocks_weights_and_returns[1:80]
eighty_stocks_returns_2019 = eighty_stocks_weights_and_returns[81:(nrow(return_matrix)+80)]

eighty_stocks_sum2019 = get_returns(return_matrix,eighty_stocks,eighty_stocks_weights)
eighty_stocks_sum2020 = get_returns(test_return_matrix,eighty_stocks,eighty_stocks_weights)
```

```{r}
ninety_stocks = get_stocks(90,corr_matrix2)

ninety_stocks_weights_and_returns = get_weights_and_returns(return_matrix,90,ninety_stocks)
ninety_stocks_weights = ninety_stocks_weights_and_returns[1:90]
ninety_stocks_returns_2019 = ninety_stocks_weights_and_returns[91:(nrow(return_matrix)+90)]

ninety_stocks_sum2019 = get_returns(return_matrix,ninety_stocks,ninety_stocks_weights)
ninety_stocks_sum2090 = get_returns(test_return_matrix,ninety_stocks,ninety_stocks_weights)
```

```{r}
entire_index = c(1:100)

entire_index_weights_and_returns = get_weights_and_returns(return_matrix,100,entire_index)
entire_index_weights = entire_index_weights_and_returns[1:100]
entire_index_returns_2019 = entire_index_weights_and_returns[101:(nrow(return_matrix)+100)]

entire_index_sum2019 = get_returns(return_matrix,entire_index,entire_index_weights)
entire_index_sum2020 = get_returns(test_return_matrix,entire_index,entire_index_weights)
```










### Example to understand weights problem
```{r}
#Solving the example formulation
objective = c(rep(1,3),0)

A = matrix(0,6,4)
b = rep(0,6)
sense = rep('>',6)

A[1,] = c(1,0,0,-1)
A[2,] = c(1,0,0,1)
b[1] = -1
b[2] = 1

A[3,] = c(0,1,0,-1)
A[4,] = c(0,1,0,1)
b[3] = -2
b[4] = 2

A[5,] = c(0,0,1,-1)
A[6,] = c(0,0,1,1)
b[5] = -3
b[6] = 3

# set up the model
model = list()
model$modelsense = 'min'
model$obj = objective
model$A = A
model$rhs = b
model$sense = sense

# solve the problem
params = list()
params$outputflag = 0
sol = gurobi(model,params=params)

# optimal allocation:
x = sol$x
```


##3 notes:

If m $\leq$ ncol (do the function)
else:
don't do it.
